---
title: 'Census 2020: Challenges for the IE'
author: "Sono Shah"
date: "2018-12-17T00:00:00"
draft: false
output: html_document
tags: ['Census']
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(readxl)
library(tidycensus)
library(hrbrthemes)
library(gghighlight)
counties <- c("Riverside County", "San Bernardino County")
dta <- read_rds("~/Dropbox/Shared/CSI/CSI Census help/census-planning-database/census_ca_tract_regions.rds")
dta <- dta %>% 
  mutate(lrs = low_response_score/100)
county_LRS <- dta %>% 
  group_by(county_name) %>%
  summarize(lrs= mean(lrs,na.rm = T),
            total_pop=sum(tot_population_acs_12_16,na.rm = T)) %>% 
  select(county_name,total_pop, lrs) %>% 
  mutate(county_color = case_when(county_name %in% counties ~ "IE", TRUE ~ "Rest of CA")) %>% 
  mutate(county_name=str_replace(county_name,"County",""))

county_LRS <- county_LRS %>% arrange(desc(lrs))
colvec <- ifelse(county_LRS$county_color =="IE", "firebrick", "gray81")
colvec <- rev(colvec)

## Reordering county_LRS$County_name
county_LRS$county_name <- factor(county_LRS$county_name, levels=c("Imperial ", "Madera ", "Merced ", "Kings ", "Tulare ", "Fresno ", "Los Angeles ", "Kern ", "San Bernardino ", "Monterey ", "Yuba ", "Mono ", "Colusa ", "Alpine ", "Del Norte ", "San Joaquin ", "San Francisco ", "Stanislaus ", "Glenn ", "Santa Barbara ", "San Benito ", "Yolo ", "Riverside ", "Mendocino ", "San Diego ", "Lake ", "Solano ", "Humboldt ", "Sacramento ", "Butte ", "Alameda ", "Lassen ", "Trinity ", "Sutter ", "Tehama ", "Santa Cruz ", "Ventura ", "Siskiyou ", "Orange ", "San Luis Obispo ", "Santa Clara ", "Napa ", "Sonoma ", "San Mateo ", "Contra Costa ", "Modoc ", "Mariposa ", "Tuolumne ", "Shasta ", "Inyo ", "Plumas ", "Sierra ", "El Dorado ", "Calaveras ", "Marin ", "Placer ", "Amador ", "Nevada "))
fct_rev(county_LRS$county_name)

```

As Census 2020 approaches, I wanted to take a closer look at the unique set of challenges that the Inland Empire faces in terms of achieving an acurate count. All the data comes from the Census planning database. In preparation for the 2020 Census, the planning database includes a census measure called `Low Response Score` or `LRS`. The `LRS` can be directly interpreted as the "predicted mail non-response rate".^[See https://www.census.gov/research/data/planning_database/2018/docs/erdman_bates_2017.pdf]

Some caveats here:

1. You are not supposed to aggregate LRS to calculate county-level estimates. See below

>The Low Response Score should NOT be aggregated (i.e., summed or averaged) to different
geographies than it is calculated. The Low Response Score is based on a linear regression model
that utilizes multiple characteristics of a given census tract or block group to predict mail nonreturn rates for that particular geography. Even though census tracts nest within a county, you
should not sum or average the Low Response Score across all census tracts in a county to arrive
at a county-level Low Response Score. Separate models would need to be developed for other
levels of geography (e.g., county, state).^[[See doc here](https://www2.census.gov/geo/pdfs/maps-data/maps/roam/ROAM_FAQ.pdf]

2. `LRS` only predicts non-reponse for MAIL
3. Subsequently, the current version of `LRS` does not account for broadband access

```{r, fig.height = 10, fig.width = 8}

county_LRS %>%
  top_n(30,wt=lrs) %>% 
  ggplot(aes(x=lrs, y= reorder(county_name,lrs))) +
  # ggplot(aes(x=Low_Response_Score, y= County_name, color =county_color )) +
  geom_point(aes(color =county_color),size=4)+
  scale_color_manual(values = c("firebrick","grey"))+
  scale_x_continuous(labels=scales::percent_format(2))+
  theme_ft_rc(grid="Y")+
  theme(axis.text.y= element_text(color=colvec),
        legend.position = "none")+
  labs(title ="County by Average LRS Score", x = "LRS SCORE", y="")
# 
# library(plotly)
# ggplotly(plot1)

```


# Estimated non-response by IE subregion

Here's a look at the distribution of tract-level non-reponse rates by tract-population. Some subregions have tracts with small populations but large non-reponse rates and others have large tract-populations with relatively small non-reponse rates.

```{r ,fig.height = 10, fig.width = 8}
dta %>%
  drop_na(region) %>%
  ggplot(aes(x=tot_population_acs_12_16, y = lrs, color = region))+
  geom_point(alpha=.6, size = 2)+
  scale_y_continuous(labels = scales::percent_format(2))+
  scale_x_continuous(labels = scales::comma)+
  labs(x = "Tract Population", y = "Low Response Score", color = "", title ="Census Tracts by Subregion")+
  theme_ft_rc(grid="")+
  theme(legend.position="none")+
  gghighlight(use_direct_label=F)+
  scale_color_ft()+
  facet_wrap(~region, nrow = 4)
```

## Total estimated non-response

```{r}
# dta %>%
  # group_by(NAME_REG) %>%
  # summarize(sum_LRS= sum(LRS_Pop, na.rm = T)) %>%
  # filter(!is.na(NAME_REG)) %>%
  # ggplot(aes(x= reorder(NAME_REG,sum_LRS), y = sum_LRS))+
  # geom_col(width=.8, fill="firebrick") +
  # coord_flip() +
  # scale_y_continuous(labels = scales::comma)+
  # labs(title= "Estimated Census Non-Response Population",
  #      subtitle="Total Population (ACS 2016 5YR) X Census LRS",
  #      x = "Total Population",
  #      y = "Census LRS", fill = "",
  #      caption="Source: Census 2018 Planning Database")+
  # theme_ipsum_rc(grid="")+
  # theme(legend.position="top")



```

